---
title: "Compute diversity from user-defined set of layers"
author: "Jean-Baptiste Féret, Florian de Boissieu"
date: "`r Sys.Date()`"
output:
  html_vignette:
    number_sections: true
vignette: >
  %\VignetteIndexEntry{CustomLayerDiversity}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

  ```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

Users may want to compute diversity from a set of raster layers differing from reflectance data.

This functionality allows for more flexible and faster application of `biodivMapR`, as all processings dedicated to the analysis of a multispectral/hyperspectral continuum (such as continuum removal, and possibly PCA if the variables supporting the analysis show minimum correlation). 

Here we will illustrate this functionality with the S2 subset used in the previous steps. 


## Computation of the spectral indices

The computation of spectral indices can be performed with the package `prosail`. Please refer to the [homepage](https://jbferet.gitlab.io/prosail/index.html) and follow instructions for installation. 

### Load additional libraries and prepare data for computation

Here, the path for the image to be processed is assumed to be defined in the `destfile` variable defined [previously](https://jbferet.github.io/biodivMapR/articles/biodivMapR_1.html){target="_blank"}.

We will first read it with the `raster` package, and get the central wavelength corresponding to each spectral band in the file from the __header file__, which is critical for the automated computation of spectral indices. 

```{r load libraries and read image}
library(prosail) # https://jbferet.gitlab.io/prosail/
library(preprocS2) # https://jbferet.gitlab.io/preprocs2/
library(stars)
library(raster)
# path (absolute or relative) for the image to process
Input_Image_File <- destfile
NameRaster <- basename(Input_Image_File)
# path (absolute or relative) for the image to process
ImBrick <- raster::brick(Input_Image_File)
HDR <- read_ENVI_header(get_HDR_name(Input_Image_File))
SensorBands <- HDR$wavelength
```


### Computation of spectral indices pre-defined in `prosail`

A set of spectral indices specific to Sentinel-2 sensors is already available. If you are not using Sentinel-2 data, then the spctral indices will still be computed, but they will use the closest band available to the Sentinel-2 bands required. 

__There are over 35 spectral indices, some showing high redundancy with each others.__

You can check the expression used to compute them [here](https://gitlab.com/jbferet/prosail/-/blob/master/R/Lib_SpectralIndices.R){target="_blank"}.

To obtain the list of spectral indices, and to compute a set of spectral indices from this list, please proceed as follows:

```{r Compute pre-defined indices}
# First define a list of spectral indices
Sel_Indices <- c('mNDVI705','CR_SWIR','LAI_SAVI')
# then compute the spectral indices
Spectral_Indices <- ComputeSpectralIndices_Raster(ImBrick, SensorBands, Sel_Indices=Sel_Indices, 
                                                  StackOut=T,ReflFactor = 10000)
# get the name of all spectral indices available
unlist(Spectral_Indices$listIndices)

# [1] "ARI1" "ARI2" "ARVI" "BAI" "BAIS2" "CHL_RE" "CRI1" "CRI2" "EVI" "EVI2" "GRVI1" "GNDVI" "IRECI" 
# [14] "LAI_SAVI" "MCARI" "mNDVI705" "MSAVI2" "MSI" "mSR705" "MTCI" "nBR_RAW" "NDI_45" "NDII" "NDVI" "NDVI_G" "NDVI705" 
# [27] "NDWI1" "NDWI2" "PSRI" "PSRI_NIR" "RE_NDVI" "RE_NDWI" "S2REP" "SAVI" "SIPI" "SR" "CR_SWIR" 
 
```

`ReflFactor` is the multiplicative factor applied on the reflectance data. Here Sentinel-2 are distributed in Int16 value types, meaning that the original reflectance was multiplied by 10000. The function then conveerts to standard reflectance values prior to computation of spectral index. 

- if `StackOut=T`: a stack of the spectral indices is directly returned in `Spectral_Indices$SpectralIndices`.

- if `StackOut=F`: a single raster layer is returned for each spectral index, accessible in `Spectral_Indices$SpectralIndices$CR_SWIR` for example.


### Computation of user-defined spectral indices through spectral index interpreter

You may want to be able to compute your own spectral index, if interested in a specific spectral index which is not in the list of pre-defined indices, or if using a sensor which is not Sentinel-2. 

The function `ComputeSpectralIndices_fromExpression` allows this flexibility, if the spectral index can be expressed as a sufficiently simple expression. Here is an illustration of the computation of the NDVI using this function: 


```{r Compute from expression}
# 1- Define the expression corresponding to the spectral index
Expression_NDVI <- '(B2-B1)/(B2+B1)'
# 2- Define the spectral bands corresponding to bands in the spectral index, identified as BXX, with XX the number of the band
Bands_NDVI <- list()
Bands_NDVI[['B1']] <- 665
Bands_NDVI[['B2']] <- 835
# 3- Compute the spectral indices
NDVI <- ComputeSpectralIndices_fromExpression(Refl = ImBrick, SensorBands = SensorBands, 
                                               ExpressIndex = Expression_NDVI, listBands = Bands_NDVI, 
                                               ReflFactor=10000, NameIndex = 'NDVI')
```

The same results should be obtained as when using pre-defined NDVI, with a correlation of 1. 

```{r Compute pre-defined NDVI}
# First define a list of spectral indices
Index <- c('NDVI')
# then compute the spectral indices
NDVI_Precomp <- ComputeSpectralIndices_Raster(ImBrick, SensorBands, Sel_Indices=Index, 
                                                  StackOut=F,ReflFactor = 10000)
cor.test(c(values(NDVI)),c(values(NDVI_Precomp$SpectralIndices$NDVI)))
```


## Stacking and writing spectral indices in format expected by biodivMapR

Once the spectral indices are computed, they will be written in __ENVI format with [BIL interleaves](https://bitbucket.org/hu-geomatics/enmap-box-idl/wiki/Data%20Format%20Definition){target="_blank"}__.

This file format is used as this image stack will directly replace the PCA file produced by biodivMapR.

It may be necessary to apply a filter on these spectral indices in order to mask outliers, as these may strongly impact the spatial variability of the features to be used, and the corresponding diversity indices. 

Here we are suggesting a method based on interquartile range (IQR), but feel free to use your own method to mask unwanted data from your stack: you can use the mask produced from the function `perform_radiometric_filtering`, if using optical data. 


### Mask data from spectral indices and apply the mask on spectral indices

```{r Produce mask based on IQR and apply it on spectral indices}
# initialize mask
Mask <- 0*Spectral_Indices$SpectralIndices$mNDVI705+1
# remove outliers from spectral indices
for (idx in Sel_Indices){
  rast <- Spectral_Indices$SpectralIndices[[idx]]
  IQRminmax <- IQR_outliers(DistVal = raster::values(rast),weightIRQ = 3)
  Mask[rast<IQRminmax[1] | rast>IQRminmax[2]] <- NA
}
# apply mask on stack and convert as stars object
StarsObj <- st_as_stars(Spectral_Indices$SpectralIndices*Mask)

```


### Save Mask file and spectral indices as raster stack


```{r Save files}
# define path where to store spectral indices (same as root path for biodivMapR output directory)
NameStack <- paste(NameRaster,'_StackIndices',sep = '')
PathIndices <- file.path(Output_Dir,NameStack)
dir.create(PathIndices,recursive = T,showWarnings = F)
# save mask in raster file
Input_Mask_File <- file.path(PathIndices,'Mask')
stars::write_stars(st_as_stars(Mask), dsn=Input_Mask_File, driver =  "ENVI",type='Byte')
# save Stack of spectral indices in raster file with spectral indices defining band names
Path_StackIndices <- file.path(PathIndices,NameStack)
write_StarsStack(StarsObj = StarsObj, dsn = Path_StackIndices, Bands=Sel_Indices, datatype='Float32')
```


## Run biodivMapR on a raster stack


### Set parameters for biodivMapR

```{r Parameterize biodivMapR}
# see additional information here
# https://jbferet.github.io/biodivMapR/articles/biodivMapR_2.html

# Define path for original image file to be processed
Input_SIstack_File <- Path_StackIndices
# trick to define the same directory as PathIndices defined in previous chunk
TypePCA <- 'noPCA'
# Define path for master output directory where files produced during the process are saved 
Output_Dir <- '../RESULTS'
# window size forcomputation of spectral diversity
window_size <- 10
# computational parameters
nbCPU <- 4
MaxRAM <- 0.5
# number of clusters (spectral species)
nbclusters <- 50

# use 10% of image to sample pixels and partition it in 20 runs of kmeans
NbPix <- sum(Mask[,,,1],na.rm = T)
nb_partitions <- 20
Pix_Per_Partition <- round(0.1*NbPix/nb_partitions)
```

### Perform Spectral species mapping

```{r Perform Spectral species mapping}
# https://jbferet.github.io/biodivMapR/articles/biodivMapR_5.html
print("MAP SPECTRAL SPECIES")
# select all spectral indices
SelectedPCs <- seq(1,dim(raster::stack(Input_SIstack_File))[3])
Kmeans_info <- map_spectral_species(Input_Image_File = Input_SIstack_File, Output_Dir = Output_Dir, SelectedPCs = SelectedPCs,
                                    PCA_Files = Input_SIstack_File, Input_Mask_File = Input_Mask_File,
                                    Pix_Per_Partition = Pix_Per_Partition, nb_partitions = nb_partitions,
                                    nbCPU = nbCPU, MaxRAM = MaxRAM, nbclusters = nbclusters, TypePCA = TypePCA)
```

### Map diversity indices

```{r Map diversity indices}
# https://jbferet.github.io/biodivMapR/articles/biodivMapR_6.html
print("MAP ALPHA DIVERSITY")
Index_Alpha <- c('Shannon')
map_alpha_div(Input_Image_File = Input_SIstack_File, Output_Dir = Output_Dir, TypePCA = TypePCA,
              window_size = window_size, nbCPU = nbCPU, MaxRAM = MaxRAM,
              Index_Alpha = Index_Alpha, nbclusters = nbclusters,LowRes = T)

print("MAP BETA DIVERSITY")
map_beta_div(Input_Image_File = Input_SIstack_File, Output_Dir = Output_Dir, TypePCA = TypePCA,
             window_size = window_size, nb_partitions=nb_partitions, nbCPU = nbCPU, MaxRAM = MaxRAM,
             nbclusters = nbclusters)

print("MAP FUNCTIONAL DIVERSITY")
# https://jbferet.github.io/biodivMapR/articles/biodivMapR_7.html
# Villeger et al, 2008 https://doi.org/10.1890/07-1206.1
map_functional_div(Original_Image_File = Input_SIstack_File, Functional_File = Input_SIstack_File,
                   Selected_Features = SelectedPCs, Output_Dir = Output_Dir,
                   window_size = window_size, nbCPU = nbCPU, MaxRAM = MaxRAM,TypePCA = TypePCA, LowRes = T)
```

### Perform validation based on a vectorized plot network

The same procedure as in the [previous tutorial](https://jbferet.github.io/biodivMapR/articles/biodivMapR_8.html){target="_blank"} can be applied.



