---
title: "Converting a raster file to the proper file format"
author: "Jean-Baptiste Féret, Florian de Boissieu"
date: "`r Sys.Date()`"
output: 
  html_vignette:
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Tutorial_raster2BIL}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

<p style="color:red">**if you are running version 1.3.0 or more recent version of __biodivMapR__ this tutorial is not needed**   </p>

__biodivMapR__ was expecting a specific raster file format as input file before version 1.3.0. 

* Raster files were expected to be raw flat-binary files. They should not contain any header, which are stored in a separate .hdr file. Hence format such as TIFF are not accepted. 
* This binary file should be written following a [Band-interleaved-by-line storage format](https://www.harrisgeospatial.com/docs/enviimagefiles.html "BIL definition").
* This file should have either `.bil` file extension, or no file extension.
* The header file should be named the same as the binary image file, but with `.hdr` file extension, and should be located in the same directory. 
* the header file should be an ASCII file and follow the [__ENVI file format__](https://www.harrisgeospatial.com/docs/ENVIHeaderFiles.html "HDR description").

A function is dedicated to conversion of a raster file into appropriate image format, named `raster2BIL`. This function is relatively straightforward to run, but users should make sure prerequisites are met before running the function.

# How to produce raster data with proper format using `raster2BIL`

`raster2BIL` can be run as follows: 

```{r raster2BIL}
library(biodivMapR)
# define input variables for raster2BIL:
# full path or relative path for the raster file to convert
Path_Raster_File <- 'Where/is/your/raster/file/?'
# you also need to define a template for the .hdr file corresponding to your sensor. 
# locate the install directory for biodivMapR
# system.file(package = "biodivMapR")`
# and write a file with description corresponding to your image (in terms of spectral bands, band name, etc)
Name_Template_HDR <- 'Name_of_hdr_template'
# you may want to convert reflectance from [0-1] real32 values to [0-10000] integer16
# see explanation below
Convert_Integer <- TRUE
Multiplying_Factor <- 10000
Multiplying_Factor_Last <- 1.0

Input_Image_File  = raster2BIL(Raster_Path = Path_Raster_File,
                               Sensor = Name_Template_HDR,
                               Output_Dir = Path_Output_Directory,
                               Convert_Integer = Convert_Integer,
                               Multiplying_Factor = Multiplying_Factor,
                               Multiplying_Factor_Last=Multiplying_Factor_Last,
                               Mask = FALSE)
```

## Input variables

The following information is accepted as input information:  

* `Raster_Path`: a string containing the full path for the raster to be converted.
* `Sensor`: a string containing the name of the sensor, which should actually be linked to a template for the corresponding header file. 
* `Output_Dir` [optional]: a string defining the path for the output directory where converted file will be stored.
* `Convert_Integer` [optional]: a boolean stating if the raster data should be converted into integer
* `Multiplying_Factor` [optional]:  a multiplying factor for the raster data. This is useful when converting real reflectance values between 0 and 1 into integer between 0 and 10000. Set `Multiplying_Factor =10000` in this case.
* `Multiplying_Factor_Last` [optional]: a multiplying factor for last band. This option is useful only if working with dry vegetation acquired with Sentinel-2 images, in order to artifically increase the reflectance of the B12 band (SWIR 2, centered at 2200 nm), so that SWIR 1 band remains below the convex hull defined during Continuum removal. __You do not need to use this option most of the time, unless you want to enhance teh sensitivity to different types of dry vegetation__.
* `Mask` [optional]: a boolean stating if the raster data corresponds to a mask (TRUE) or not (FALSE). If it is a mask, masked pixels should have a value of `0`, and pixels to keep for the analysis should have a value of `1`. raster2BIL converts image information into unsigned 8 bits integer.


Please note that 
i) the coordinate system and minimal metadata (spatial resolution, geographic information...) are expected to be provided in the original image or image + header defined in `Raster_Path` if user want them to be included in the final image products.

ii) `Sensor` should refer to the name of a `.hdr` file stored in the installation directory of `biodivMapR`
Please type `system.file(package = "biodivMapR")` in order to get the location of the install directory and create a proper `.hdr` file to be stored in `extdata/HDR/`, containing proper spectral bands __defined in nanometers__ if using optical data and using `perform_radiometric_filtering` and/or `Continuum_Removal = TRUE` in your process. Make sure that the __spectral bands in the binary file are stored with wavelengths following ascending order__. 

iii) `Convert_Integer=TRUE` saves 50% space if original image is stored in real values (32 bits per value). However, user need to make sure hat their original data are compatible with conversion: if reflectance values are stored as  real value between 0 and 1, the final image file will be unusable, unless proper multiplying factor is applied to the full image data (for example `Multiplying_Factor =10000`). Sentinel-2 reflectance data downloaded from ESA hub are stored as integer (16 bits) values between 0 and 10 000 instead of real (32 bits) values between 0 and 1, so integer conversion is unnecessary.

## output variable

`raster2BIL` returns a string containing the full path for the raster once converted.

You can directly use this output variable as input for the next steps of the process.
