---
title: "Perform spatial & spectral filtering"
author: "Jean-Baptiste Féret, Florian de Boissieu"
date: "`r Sys.Date()`"
output: 
  html_vignette:
    number_sections: true
vignette: >
  %\VignetteIndexEntry{DataFiltering}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

### Produce a vegetation / cloud / shade mask

The full radiometric filtering can be peformed with a unique function, including: 

* NDVI thresholding

* NIR thresholding

* Blue thresholding

The spectral bands corresponding to the Blue, Green and NIR spectral bands can be defined as optional parameters. 

```{r perform radiometric filtering}
print("PERFORM RADIOMETRIC FILTERING")
Input_Mask_File <- perform_radiometric_filtering(Image_Path = Input_Image_File, Mask_Path = Input_Mask_File,
                                                 Output_Dir = Output_Dir, TypePCA = TypePCA,
                                                 NDVI_Thresh = NDVI_Thresh, Blue_Thresh = Blue_Thresh,
                                                 NIR_Thresh = NIR_Thresh)
```


### Produce a spectral filter

Spectral bands or spectral domains can be discarded during the process. This is particularly useful when spectral domains with high SNR are includd in the image. 

By default, the spectral domains corresponding to atmospheric water absorption are masked. 

However, these spectral domains can be extended, or included with the variable `Excluded_WL`, which is an input variable for the function `perform_PCA`. The default values are the following for the spectral domains to exclude (in nanometers):

```{r Excluded_WL}
Excluded_WL <- c(0, 400)
Excluded_WL <- rbind(Excluded_WL, c(895, 1005))
Excluded_WL <- rbind(Excluded_WL, c(1180, 1480))
Excluded_WL <- rbind(Excluded_WL, c(1780, 2040))
```

<br/><br/>

The PCA transformation and dimensional reduction are performed in the [next step](https://jbferet.github.io/biodivMapR/articles/biodivMapR_4.html){target="_blank"}.
