---
title: "Functional Diversity"
author: "Jean-Baptiste Féret, Florian de Boissieu"
date: "`r Sys.Date()`"
output:
  html_vignette:
    number_sections: true
vignette: >
  %\VignetteIndexEntry{FunctionalDiversity}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```


The functional diversity metrics (Richness, Evenness, Divergence) can be computed directly after dimensionality reduction. Here, we assume that each selected component correspond to a trait value (each trait is scaled over the full image), and each pixel corresponds to a species/individual (no clustering is performed, so one pixel from an individual spatial unit is seen as a different species).

```{r Functional diversity metrics}
################################################################################
##              MAP FUNCTIONAL DIVERSITY METRICS FRic, FEve, FDiv             ##
##          (Villeger et al, 2008 https://doi.org/10.1890/07-1206.1)          ##
################################################################################
## read selected features from dimensionality reduction
Selected_Features <- read.table(Sel_PC)[[1]]
## path for selected components
map_functional_div(Original_Image_File = Input_Image_File, Functional_File = PCA_Files,
                   Selected_Features = Selected_Features, Output_Dir = Output_Dir,
                   window_size = window_size, nbCPU = nbCPU, MaxRAM = MaxRAM,TypePCA = TypePCA)
```

These diversity metrics can also be computed directly from plots defined as polygons in the extent of the rasters. This step is key for validation, and is explained in the [next step](https://jbferet.github.io/biodivMapR/articles/biodivMapR_8.html){target="_blank"}.
