---
title: "Spectral species mapping"
author: "Jean-Baptiste Féret, Florian de Boissieu"
date: "`r Sys.Date()`"
output:
  html_vignette:
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Spectralspecies}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```


The first step towards $\alpha$ and $\beta$ diversity mapping corresponds to the computation of a `SpectralSpecies` map, which identifies the cluster ('spectral species') assigned to each pixel in the image, after k-means clustering is performed. Most of the input parameters are obtained when running `perform_PCA`.

Info about K-means clustering and the path for the spectral species file can be obtained as outputs. 


```{r Spectral species map}
print("MAP SPECTRAL SPECIES")
Kmeans_info <- map_spectral_species(Input_Image_File = Input_Image_File, Output_Dir = Output_Dir,
                                    PCA_Files = PCA_Files, Input_Mask_File = Input_Mask_File,
                                    Pix_Per_Partition = Pix_Per_Partition, nb_partitions = nb_partitions,
                                    nbCPU = nbCPU, MaxRAM = MaxRAM, nbclusters = nbclusters, TypePCA = TypePCA)
```

SpectralSpecies is then stored in a raster file located here:

```{r Spectral Species Path}
Kmeans_info$SpectralSpecies
'RESULTS/S2A_T33NUD_20180104_Subset/SPCA/SpectralSpecies'

```



$\alpha$ and $\beta$ diversity maps can then be computed based on this `SpectralSpecies` raster.

The computation of $\alpha$ and $\beta$ diversity maps is performed in the [next step](https://jbferet.github.io/biodivMapR/articles/biodivMapR_06.html){target="_blank"}.
