---
title: "`biodivMapR` applied with biophysical trait estimation with `prosail`"
author: "Jean-Baptiste Féret, Florian de Boissieu"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
  html_vignette:
    number_sections: true
vignette: >
  %\VignetteIndexEntry{biodivMapR and prosail}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

This tutorial will guide you through the application of `prosail` in order to 
produce biophysical properties from hybrid inversion, then used to estimate 
spectral diversity with `biodivMapR`.

This is the tutorial associated to the book chapter 
[`A Plant Trait-Based Approach to Map Forest Biodiversity Using Imaging Spectroscopy and Physical Modeling`](https://doi.org/10.1007/978-3-031-99665-8_7), which is part of the 
book [`R Coding for Ecology`](https://link.springer.com/book/10.1007/978-3-031-99665-8), 
in the **Use R!** book series (Springer).


**Warning**: this tutorial is appropriate with prosail v3.0.0 and later.
please install the 'dev_joss' branch if the master branch is <3.0.0

```{r, install prosail}
devtools::install_gitlab('jbferet/prosail@dev_joss')
```



## download imaging spectroscopy data required for the tutorial

The image can be downloaded from [this webpage](https://doi.org/10.57745/SDPRYY).
Three files are available on the webpage:

- `Readme_FAbas_HS_mosaic-2.pdf`

- `Fabas_HS_mosaic.hdr`: header file for the image data, contains all metadata 
such as central wavelengths of the different bands, image size, projection...

- `Fabas_HS_mosaic.tif`: the binary image. This is a surface reflectance image, 
which is geometrically and atmospherically corrected. 
Image spatial resolution is 4m. a multiplying factor of 10000 was applied on 
reflectance (10000 = 100% reflectance).

The data should be placed in a directory defined as `data_dir` in the next code 
chunk.

Figure 1 shows an RGB representation of the imaging spectroscopy data acquired 
over the Fabas forest.

```{r, eval=TRUE, echo = FALSE, out.width="100%"}
url <- "https://gitlab.com/jbferet/myshareddata/-/raw/master/biodivMapR_vignette_illustrations/05_rgb_hyspex.png"
knitr::include_graphics(url)
```

<center>
  Fig. 1. color composition of Hyspex imaging spectroscopy acquisition. 
  
  R = 661 nm, G = 559 nm, B = 458 nm. 
</center> 
<p>&nbsp;</p>


## define input / output directories

The input data directory is defined, as well as the directory where vegetation
biophysical parameters obtained from `prosail` hybrid inversion. 
This tutorial uses functions performing raster tiling when processing `prosail` 
hybrid inversion, hence the definition of a subdirectory named `tiles`.

```{r, eval=FALSE}
# clean workspace
rm(list = ls(all=TRUE)); gc()
if (rstudioapi::isAvailable()) 
  setwd(dirname(rstudioapi::getSourceEditorContext()$path))
library(prosail)

# 1.1- define input & output directories
data_dir <- './01_DATA/fabas'
prosail_dir <- './03_RESULTS/fabas/prosail_inversion'
prosail_tiles <- file.path(prosail_dir, 'tiles')
dir.create(prosail_tiles, showWarnings = F, recursive = T)
```

## prepare imaging spectroscopy data for processing: spectral/spatial masking

### get metadata and additional information

Identify the imaging spectroscopy raster, and get metadata from the header file. 
The central wavelength and full width at half maximum (FHWM) are needed to 
produce a spectral response function which will then be needed to simulate
surface reflectance with sensor response. 

```{r, eval=FALSE}
# 1.2- define input & output directories
input_raster_path <- file.path(data_dir,'Fabas_HS_mosaic.tif')

# get information on wavelength and fwhm from header
hdr <- read_envi_header(hdr_path = get_hdr_name(image_path = input_raster_path))
sensor_bands <- hdr$wavelength
sensor_fwhm <- hdr$fwhm
```



### filter out shaded, non vegetated, and hazy pixels. 

A basic filtering process can be applied in order to mask irrelevant pixels 
(no vegetation) and pixels with poor quality (hazy/cloudy or shaded areas).
The function `radiometric_filtering` implemented in the package 
[`biodivMapR`](https://jbferet.github.io/biodivMapR/index.html) performs a 
filtering combining multiple criteria. 

- non vegetated pixels are filtered out with a NDVI thresholding: all pixels with 
`NDVI < NDVI_Thresh` are masked.

- hazy / cloudy pixels are filtered out with a Blue band thresholding: all pixels
with `R\text{blue} > Blue_Thresh` are masked.

- shaded pixels are filtered out with a near infrared band thresholding: all 
pixels with `R\text{nir} < NIR_Thresh` are masked.


```{r, eval=FALSE}
##  1.3- spatial filtering: discard with hazy/cloudy, shaded & no vegetation
library(biodivMapR)
NDVI_Thresh <- 0.5
Blue_Thresh <- 600
NIR_Thresh <- 1500
mask_path <- radiometric_filtering(input_raster_path = input_raster_path, 
                                   input_rast_wl = sensor_bands,
                                   output_dir = prosail_dir, 
                                   NDVI_Thresh = NDVI_Thresh, 
                                   Blue_Thresh = Blue_Thresh, 
                                   NIR_Thresh = NIR_Thresh)
```

If no mask available, then `mask_path` should not be specified as input for
`apply_prosail_inversion`.

### filter out spectral bands with low signal to noise ratio (SNR)

Additional information corresponding to atmospheric transmittance are also 
downloaded in order to perform spectral filtering for low transmittance. 
Low SNR is **arbitrarily** defined as a threshold of 0.85. 
Please investigate how to filter out spectral domains with low SNR on your own 
imaging spectroscopy data as it may depend on sensor, conditions of acquisitions, 
and other factors. 
An additional band showing strong noise after careful analysis is also discarded. 


```{r, eval=FALSE}
# 1.4- spectral filtering
# atmospheric transmittance file available on a repository
git_repository <- 
  'https://gitlab.com/jbferet/myshareddata/-/raw/master/Fabas_HS_AtmoT'
git_file <- 'AtmoT_Fabas_Hyspex.csv'
url <- file.path(git_repository, git_file)
path_atmo <- file.path(data_dir,git_file)
download.file(url = url, destfile = path_atmo)

# spectral filtering #1: discard bands with transmission factor<0.85
transmittance <- read.delim(path_atmo, header = TRUE)
low_trans <- which(transmittance$Transmittance<0.85)
# spectral filtering #2: discard noisy band remaining
noise <- which((sensor_bands > 2200 & sensor_bands < 2205))
# combine bands to discard
wl_elim <- c(low_trans, noise)
```


## perform hybrid inversion of `prosail` on imaging spectroscopy data

The hybrid inversion implemented in the `prosail` package is a bagging 
prediction of biophysical properties: a LUT is simulated with prosail and 
resampled in order to produce multiple datasets including a limited number of 
samples. 
Then a set of individual support vector regression (SVR) models is trained from 
each reduced dataset.

The hybrid inversion consists in two main stages for users: 

- training of a machine learning regression model (SVR) for each biophysical 
property of interest based on surface reflectance factor LUT simulated with 
PROSAIL

- application of the SVR models to a dataset provided as a matrix or as a raster



### preparation for a training dataset

The hybrid inversion can be easily parameterized to adapt to any type of sensor, 
input parameter distribution, type of noise. 

alternative machine learning (ML) methods will be implemented in the future, and 
suggestions are welcome for the implementation of alternative algorithms.
Let's go. 

#### biophysical variables of interest

`prosail` takes into account the influence of multiple biophysical properties. 
Each input parameter can be theoretically estimated through hybrid inversion. 
However, the relevance and uncertainty associated with these estimates 
strongly depends on i) the type of sensor (spectral information provided), 
ii) the signal to noise ratio of the data, and iii) the type of vegetation 
(keeping in mind that `prosail` is based on a turbid hypothesis, hence, not 
appropriate for heterogeneous forests).


Here, we will estimate four biophysical properties which will then be used to 
compute spectral diversity. 
These are: 
- leaf area index (`lai`)
- leaf chlorophyll content (`chl`)
- equivalent water thickness (`ewt`)
- leaf mass per area  (`lma`)

```{r, eval=FALSE}
# biophysical variables to estimate
parms_to_estimate <- c('lai', 'chl', 'ewt', 'lma')
```

#### spectral domain of inversion

The accuracy of the estimation of vegetation biophysical properties from 
remotely sensed data varies with the spectral information provided to the ML
algorithm. 
Here, we arbitrarily define spectral domains for each biophysical property of 
interest. 
Spectral feature selection may be used in order to refine the optimal spectral 
domain for each property.
Noise level to be applied on simulated data is also defined for each property. 


```{r, eval=FALSE}
# spectral domain used for estimation of each variable, discard wl_elim
selected_bands <- noise_level <- list()
wl_train <- data.frame(min = c('lai'=700, 'chl'=700, 'ewt'=1200, 'lma'=1200),
                       max = c('lai'=900, 'chl'=900, 'ewt'=2400, 'lma'=2400))
for (parm in parms_to_estimate) {
  bandsel <- which(sensor_bands > wl_train[parm, 'min'] & 
                     sensor_bands < wl_train[parm, 'max'])
  selected_bands[[parm]] <- sensor_bands[bandsel[-na.exclude(match(wl_elim, bandsel))]]
  # 2% relative gaussian noise applied to reflectance data
  noise_level[[parm]] <- 0.02        
}
```


#### geometry of acquisition

A rough estimate of the valid range for the geometry of acquisition 
should be provided, ideally. 

- observer zenith angle `tto`

- sun zenith angle `tts`

- relative azimuth angle `psi`

```{r, eval=FALSE}
# define geometry of acquisition based on location and time of acquisition
geom_acq <- data.frame(min = c('tto' = 0, 'tts' = 50, 'psi' = 0), 
                       max = c('tto' = 15, 'tts' = 70, 'psi' = 360))
```

#### input parameter distribution

Users can define distribution and range for each input parameter of `prospect`. 
However, for the sake of straightforwardness, the parameterization defined in 
the [Algorithm Theoretical Based Document](https://step.esa.int/docs/extra/ATBD_S2ToolBox_V2.1.pdf) 
(ATBD) of the Sentinel-2 toolbox will be used here.


```{r, eval=FALSE}
# use same distribution as used in ATBD for Sentinel-2 data
input_prosail <- get_input_prosail(atbd = TRUE, geom_acq = geom_acq)
# waiting for updated soil database, deactivate soil v3 
input_prosail$soil_brightness <- input_prosail$soil_ID  <- NULL
```


#### spectral response function

The spectral response function (SRF) needs to be defined in order to simulate sensor 
reflectance. 

The SRF of several sensors is already implemented in `prosail`: Sentinel-2, 
Landsat-7, Landsat-8, Landsat-9, SPOT 6/7, MODIS, Pleiades, and Venus. 

Here, the SRF of the hyperspectral sensor is produced from central wavelengths 
and FWHM based on a gaussian response model. 


```{r, eval=FALSE}
# get sensor response function corresponding to the sensor
sensor_name <- 'Hyspex'
srf <- get_srf_sensor(sensor_name = sensor_name, 
                      wl = sensor_bands, 
                      fwhm = sensor_fwhm)
```


### train ML regression algorithms with simulated data

The function `train_prosail_inversion` performs the training of the ML algorithm 
based on the information generated in the previous steps. 
It provides a set of regression models for each biophysical property to estimate. 

```{r, eval=FALSE}
message('training ML regression with prosail simulations')
options <- set_options_prosail(fun = 'train_prosail_inversion')
options$noise_level <- noise_level
hybrid_model <- train_prosail_inversion(input_prosail = input_prosail, 
                                        parms_to_estimate = parms_to_estimate,
                                        geom_acq = geom_acq, 
                                        srf = srf,
                                        selected_bands = selected_bands,
                                        output_dir = prosail_dir, 
                                        options = options)
```

### apply ML regression algorithms on images

The function `apply_prosail_inversion` applies ML regression trained at the 
previous step on raster data. 
As for [spectral indices](https://jbferet.gitlab.io/tutorial_puertorico2025/compute-spectral-indices-from-sentinel-2-image.html#computing-spectral-indices-from-sentinel-2-raster), a grid is first defined 
on the aoi. 
Model inversion is then applied individually for each tile. 
The current version of prosail does not support multiprocessing for the 
application of ML algorithm, but suggestions are welcome for its implementation. 


```{r, eval=FALSE}
# apply models on image
message('applying ML regression on image')
options <- set_options_prosail(fun = 'apply_prosail_inversion_per_tile')
options$tiling <- TRUE      # process image per square tile
options$tile_size <- 1000   # dimensions of individual square tile (in meters)
BPvars <- apply_prosail_inversion(raster_path = input_raster_path, 
                                  mask_path = mask_path, 
                                  output_dir = prosail_tiles,
                                  hybrid_model = hybrid_model, 
                                  band_names = sensor_bands,
                                  selected_bands = selected_bands, 
                                  options = options)
```

### resulting biophysical properties

The maps corresponding to the biophysical variables estimated from imaging 
spectroscopy data are displayed below. 

```{r, eval=TRUE, echo = FALSE, out.width="100%"}
url <- "https://gitlab.com/jbferet/myshareddata/-/raw/master/biodivMapR_vignette_illustrations/05_bp_prosail.png"
knitr::include_graphics(url)
```

<center>
  Fig. 2. maps of vegetation biophysical variables  `chl`, `lai`, `ewt` and 
  `lma` obtained over the Fabas forest. 
  The variables were estimated from physical model hybrid inversion applied to 
  imaging spectroscopy and performed with the `prosail` package.
</center> 
<p>&nbsp;</p>



## apply `biodivMapR` on biophysical prioperties estimated with `prosail`

### parameterize `biodivMapR`

The directory where tiles corresponding to biophysical properties are stored 
needs to be provided. 
`biodivMapR` will then identify files with pattern matching with each 
biophysical property, and the identifier of the tile. 
the grid produced during the `prosail` inversion is also used

Multiple options can be defined with the `set_options_biodivMapR` function. 
Here, we defined `options$moving_window <- TRUE`. 
This results in a moving window process instead of cell grid process, with map 
produced with the same spatial resolution as original input data. 
While this option produces visually more satisfying outputs, it results in 
signifiantly higher computation time. 
Hence, **we do not recommend setting this option to TRUE if large image datasets are processed**.


```{r parametrize biodivMapR}
# -----------------------------------------------------------------------------
# prepare for biodivMapR to run on the set of biophysical properties estimated 
# in previous step
# -----------------------------------------------------------------------------
# define output directories
prosail_dir <- file.path(output_dir, 'prosail_inversion/tiles/mean')

# directory where mask is stored
mask_dir <- file.path(output_dir, 'mask_tiles')
dir.create(path = mask_dir, showWarnings = FALSE, recursive = TRUE)

# directory where diversity maps are stored
biodivMapR_dir <- file.path(output_dir, 'biodivMapR')
dir.create(biodivMapR_dir, showWarnings = F, recursive = T)

# define grid to perform diversity mapping
aoi_path <- file.path(input_raster_dir, 'aoi_tiles_1000m.gpkg')
aoi <- sf::st_read(dsn = aoi_path, quiet = TRUE)
nb_plots <- length(aoi$geom)
plots <- get_plot_list(dsn = aoi_path, nbdigits = nchar(nb_plots))

# -----------------------------------------------------------------------------
# run biodivMapR 
# -----------------------------------------------------------------------------
# define biodivMapR parameters and options
window_size <- 10
nbCPU <- 6

# set optional parameters
options <- set_options_biodivMapR(fun = 'biodivMapR_full_tiles', options = NULL)
options$alpha_metrics <- c('hill', 'shannon') 
options$beta_metrics <- TRUE
options$moving_window <- TRUE
options$nb_clusters <- 25
site_name <- 'Fabas_HS_mosaic'

```


### run `biodivMapR`

`biodivMapR` can run directly using a unique function `biodivMapR_full_tiles`, 
as illustrated in the code hereafter. 

```{r run biodivMapR}
# run biodivmapR on tiles defined when processing prosail inversion
ResDiv <- biodivMapR_full_tiles(feature_dir = prosail_dir, 
                                list_features = parms_to_estimate, 
                                site_name = site_name, 
                                mask_dir = mask_dir, 
                                plots = plots, 
                                output_dir = biodivMapR_dir, 
                                window_size = window_size, 
                                nbCPU = nbCPU, 
                                options = options)


```


Figure 3 displays the Hill index obtained from `biodivMapR`, and allows 
comparison with the forest stand composition corresponding to pure (blue) and 
mixed (red) forest stands, as described in the [BD Foret](https://geoservices.ign.fr/bdforet).

```{r, eval=TRUE, echo = FALSE, out.width="100%"}
url <- "https://gitlab.com/jbferet/myshareddata/-/raw/master/biodivMapR_vignette_illustrations/05_alpha_ground.gif"
knitr::include_graphics(url)
```

<center>
  Fig. 3. map of of $\alpha$-diversity (hill index) obtained from `biodivMapR` 
  analysis on biophysical variables from imaging spectroscopy acquisition.
  The variables were estimated from physical model hybrid inversion applied to 
  imaging spectroscopy and performed with the `prosail` package.
  The map is compared with forest stand composition corresponding to pure (blue)
  and mixed (red) forest stands.
</center> 
<p>&nbsp;</p>




Figure 4 displays the $\beta$-diversity obtained from `biodivMapR`, and allows 
comparison with the forest stand composition, as described in the 
[BD Foret](https://geoservices.ign.fr/bdforet).

```{r, eval=TRUE, echo = FALSE, out.width="100%"}
url <- "https://gitlab.com/jbferet/myshareddata/-/raw/master/biodivMapR_vignette_illustrations/05_beta_ground.gif"
knitr::include_graphics(url)
```

<center>
  Fig. 4. map of of $\beta$-diversity obtained from `biodivMapR` analysis on 
  biophysical variables from imaging spectroscopy acquisition.
  The variables were estimated from physical model hybrid inversion applied to 
  imaging spectroscopy and performed with the `prosail` package.
  The map is compared with forest stand composition corresponding to pure (blue)
  and mixed (red) forest stands.
</center> 
<p>&nbsp;</p>


### validate `biodivMapR` with ground data

The diversity metrics computed with `biodivMapR` can then be compared with 
ground information obtained from forest inventories.

The validation plots & inventories can be downloaded from a `gpkg` file 
available at [this webpage](https://doi.org/10.57745/NHRNOR).

The code below allows arranging the ground information. 

```{r validate biodivMapR}

# -----------------------------------------------------------------------------
# prepare inventory data for analysis and comparison with biodivMapR estimates
# -----------------------------------------------------------------------------
# define input & output directories
input_vect_dir <- '../01_data/vector_data'
prosail_bp_vrt_dir <- file.path(output_dir, 'prosail_inversion/tiles/vrt')
mask_dir <- file.path(output_dir, 'mask_tiles')
biodivMapR_dir <- file.path(output_dir, 'biodivMapR')
validation_dir <- file.path(output_dir, 'validation')
dir.create(path = validation_dir, showWarnings = FALSE, recursive = TRUE)

# produce virtual rasters for spectral features to be used
# create a vrt
bp_path <- list()
for (bp in parms_to_estimate)
  bp_path[[bp]] <- list.files(path = prosail_bp_vrt_dir, pattern = bp, full.names = T)

# mask create a vrt
listfiles <- list.files(path = mask_dir, full.names = T)
listfiles <- listfiles[grep(pattern = '.tiff', x = listfiles)]
mask_path <- file.path(validation_dir, paste0(site_name, '_mask_mosaic.vrt'))
v <- terra::vrt(x = listfiles, filename = mask_path, overwrite = TRUE)

# arrange inventory data to get diversity metrics
# The inventory data needs a bit of process to get the abundance table which 
# will then be used to produce diversity metrics.
# The function get_inventory_from_species_list will to the job.
get_inventory_from_species_list <- function(validation_plots){
  SpList <- unique(unlist(validation_plots[[c('Species1', 'Species2', 
                                              'Species3', 'Species4', 
                                              'Species5')]]))
  SpList <- sort(na.omit(SpList))
  # create species abundance table from inventory data
  inventory <- data.frame(matrix(data = 0,  
                                 nrow = length(validation_plots$ID), 
                                 ncol = length(SpList)))
  colnames(inventory) <- SpList
  rownames(inventory) <- validation_plots$ID
  for (PlotID in validation_plots$ID){
    whichPlot <- which(validation_plots$ID == PlotID)
    for (sp in seq_len(5)){
      wsp <- which(SpList == validation_plots[[paste0('Species',sp)]][[1]][whichPlot])
      if (length(wsp)>0){
        Abundspi <- paste0('%Cover_Species',sp)
        inventory[whichPlot,SpList[wsp]] <- validation_plots[[Abundspi]][[1]][whichPlot]
      }
    }
  }
  return(inventory)
}

# We can then get the diversity metrics we are interested in.
vect_path <- file.path(input_vect_dir, 'forestPLOT_Fabas_2019.gpkg')
validation_plots <- terra::vect(vect_path)
inventory <- get_inventory_from_species_list(validation_plots)
# compute diversity metrics from abundance table
plotID <- validation_plots$ID
shannon_obs <- vegan::diversity(inventory, index = "shannon")
richness_obs <- vegan::specnumber(inventory)
bc_obs <- vegan::vegdist(inventory, method = "bray")


# extract spectral diversity metrics from plots footprint
# We can then use information produced at the previous step in order to get 
# diversity metrics from plot footprint. 
# plots are expected to be polygons, not points. Apply buffer if needed
load(file.path(biodivMapR_dir, 'Kmeans_info.RData'))
load(file.path(biodivMapR_dir, 'Beta_info.RData'))
input_rast <- lapply(X = bp_path, FUN = terra::rast)
input_mask <- terra::rast(mask_path)
alpha_metrics <- c('richness', 'shannon', 'simpson', 'hill') 
fd_metrics <- c('FRic', 'FEve', 'FDiv', 'FDis', 'FRaoq') 
validation <- get_diversity_from_plots(input_rast = input_rast, 
                                       input_mask = input_mask, 
                                       validation_vect = validation_plots, 
                                       Kmeans_info = Kmeans_info, 
                                       Beta_info = Beta_info, 
                                       alpha_metrics = alpha_metrics,
                                       fd_metrics = fd_metrics)

```


### compare `biodivMapR` with ground data

The following code allows production of a figure comparing measured and 
estimated diversity metrics.

```{r produce plots}

# produce scatterplots comparing inventory data with spectral diversity
# The code below produces a scatterplot comparing diversity metrics derived from
# forest inventories with diversity metrics produced with biodivMapR.

# richness
predicted <- validation$specdiv$richness_mean
df_richness <- data.frame('measured' = richness_obs,
                          'predicted' = predicted)
# shannon
predicted <- validation$specdiv$shannon_mean
df_shannon <- data.frame('measured' = shannon_obs,
                         'predicted' = predicted)
# Bray-Curtis dissimilarity
BCdissmat <- c(as.dist(validation$BC_dissimilarity, diag = FALSE))
Field_Beta <- c(as.dist(bc_obs, diag = FALSE))
df_beta_estimated <- data.frame('measured' = Field_Beta,
                                'predicted' = BCdissmat)

plot_data <- data.table::rbindlist(list(
  `Spectral species richness` = df_richness,
  `Shannon's H` = df_shannon,
  `BC dissimilarity` = df_beta_estimated), idcol = 'Metric')
plot_data <- dplyr::mutate(plot_data, 
                           Metric = factor(Metric, 
                                           levels = unique(Metric), 
                                           labels = unique(Metric)))

plotval <- ggplot(plot_data, aes(measured, predicted, color=Metric))+
  geom_point()+
  geom_smooth(method="lm", color="black")+
  ggplot2::facet_wrap(~Metric, scales="free", ncol=3)+
  ggpmisc::stat_correlation(method = "spearman", 
                            label.x = "right", 
                            label.y = "bottom", 
                            color="black")+
  guides(color="none") +
  annotate("label", colour = "black", size = 10)

filename <- file.path(validation_dir, 'Validation_AlphaBeta.png')
ggsave(filename, plot = plotval, device = 'png', path = NULL, scale = 0.75, 
       width = 12, height = 4, units = "in", dpi = 600, limitsize = TRUE)
```




Figure 5 displays the comparison between diversity metrics obtained from 
forest inventories and those obtained from `biodivMapR` applied to biophysical 
properties obtained from imaging spectroscopy.

```{r, eval=TRUE, echo = FALSE, out.width="100%"}
url <- "https://gitlab.com/jbferet/myshareddata/-/raw/master/biodivMapR_vignette_illustrations/05_Validation_AlphaBeta.png"
knitr::include_graphics(url)
```

<center>
  Fig. 5. comparison between diversity metrics obtained from forest inventories 
  and those obtained from `biodivMapR` applied to biophysical 
  properties obtained from imaging spectroscopy.
  from left to right: species vs. spectral richness; species vs. spectral 
  Shannon, species vs. spectral bray curtis dissimilarity.
</center> 
<p>&nbsp;</p>
